---
title: "Session 1 Lesson Plan"
author: "Myfanwy Johnston"
date: "January 20, 2021"
output: 
  html_document:
    theme: lumen
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

Our ultimate goal here is that by the end of the _second_ training session, you will understand how to make the reproducible report,  and how to update it with new data in accordance with best practices for reproducible workflows.

So, in this first training session, we will be making a practice reproducible report called “Cod Report”.  “Cod Report” will consist of two different .Rmd documents.  Its setup and organization will follow best practices for reproducibility, using a project-oriented workflow.  The first three exercises below (I-III) are designed to mirror the process of creating a more complicated reproducible report from start to finish. The last exercise sets us up for the second training session.


## Exercise 1

Set up a local project directory in accordance with best practices for project-oriented workflow

1. Create project folder “codfish” somewhere on your hard drive.  
    - This is your project directory.
2. Within the project directory, create project subfolders: 
    - /R, /data_raw, /data_clean, /docs
3. Move or copy “CodParasite.csv” into the /data_raw subfolder
4. Open RStudio, select “New Project” and using the “Existing Directory” option, navigate to your codfish project folder
5. Create a new R script and save it as “utils.R” to the codfish project’s R subfolder
    - Save some basic functions to utils.R, including the function “condition” that we wrote in the R course.
    - You can find the code for a condition function [here](https://gist.github.com/Myfanwy/2445f0972f0f6a65d75a0750fdecf779)
6. Create another R script called “clean_codparasite.R”.  In it:
    - read CodParasite.csv into R
    - source “utils.R”
    - perform basic cleaning operations on the data
    - remove NAs in columns of interest
    - add a new column for fish condition
    - subset down to the columns we want
    - save final object “codparasite_clean.rds” to /data_clean/
    - Comment at the top that the output is “data_clean/codparasite_clean.rds”


----- Rmd Introductory slides -----


## Exercise 2

Create a functioning (knits without error) .Rmd document/report


1. In your codfish project, create a new Q1_tables.Rmd and save it to /docs
2. In the setup chunk, load “codparasite_clean.rds” and comment on where it was created
3. Create two more chunks, one for table 1 and one table 2.  
    - Create the tables using knitr::kable().
    - Set your knit directory to project folder
4. Knit the .Rmd to Word.
5. Create a new Q2_plots.Rmd, similar to Q1_tables but with plots instead of tables, and save it to /docs.
    - make sure the setup chunk is named something different than what it’s named in the Q1 Rmd
6. Create a new CodReport.Rmd
7. Outline the two sections in markdown
8. Beneath each section, insert a named code chunk and use the child= argument to name the other Rmds that will be sourced there
9. Click “knit”


## Exercise 3

Install CDECRetrieve, fishpals, and tagtales packages from GitHub in preparation for next week.

1. Install the `remotes` package:

```{r, eval=FALSE, echo=TRUE}
install.packages("remotes")
```

2. Install each subsequent package using `remotes::install_github()`
    - If you get a message about updating packages, enter 1 into the console to update all

```{r, eval=FALSE, echo=TRUE}
remotes::install_github("flowwest/CDECRetrieve")
remotes::install_github("fishsciences/fishpals")
remotes::install_github("Myfanwy/tagtales")
```

